Diffantom: whole-brain diffusion MRI phantoms derived from real datasets of the \acrlong{hcp}

### 2015-12-22 paper/abbreviations.tex
final version for bioRxiv
reply reviewers' comments

% -*- root: main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-06-16 15:45:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-08-14 09:16:55

\newacronym{dmri}{dMRI}{diffusion MRI}
\newacronym{fa}{FA}{fractional anisotropy}
\newacronym{adc}{ADC}{anisotropic diffusion coefficient}
% \newacronym{hcp}{HCP}{Human Connectome Project}
\newacronym{fod}{FOD}{fiber orientation distribution}
\newacronym{wm}{WM}{white matter}
\newacronym{gm}{GM}{gray matter}
\newacronym{cgm}{cGM}{cortical gray matter}
\newacronym{dgm}{dGM}{deep gray matter}
\newacronym{csf}{CSF}{cerebrospinal fluid}
\newacronym{snr}{SNR}{signal-to-noise ratio}
\newacronym{csd}{CSD}{constrained spherical deconvolution}
\newacronym{cc0}{CC0}{Creative Commons Zero licence}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{pve}{PVE}{partial volume effect}
\newacronym{cst}{CST}{corticospinal tract}
\newacronym{t1}{T1w}{T1-weighted}

@@

>>>>

% -*- root: main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-06-16 15:45:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-12-16 16:39:23

\newacronym{dmri}{dMRI}{diffusion MRI}
\newacronym{fa}{FA}{fractional anisotropy}
\newacronym{adc}{ADC}{anisotropic diffusion coefficient}
% \newacronym{hcp}{HCP}{Human Connectome Project}
\newacronym{fod}{FOD}{fiber orientation distribution}
\newacronym{fodf}{fODF}{fiber orientation distribution function}
\newacronym{wm}{WM}{white matter}
\newacronym{gm}{GM}{gray matter}
\newacronym{cgm}{cGM}{cortical gray matter}
\newacronym{dgm}{dGM}{deep gray matter}
\newacronym{csf}{CSF}{cerebrospinal fluid}
\newacronym{snr}{SNR}{signal-to-noise ratio}
\newacronym{csd}{CSD}{constrained spherical deconvolution}
\newacronym{cc0}{CC0}{Creative Commons Zero licence}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{pve}{PVE}{partial volume effect}
\newacronym{cst}{CST}{corticospinal tract}
\newacronym{t1}{T1w}{T1-weighted}

@@

\newglossaryentry{bids}%
{
	name={BIDS},
	first={BIDS standard \cite[Brain Imaging Data Structure,][]{gorgolewski_brain_2015}},
	long={Brain Imaging Data Structure},
	description={Brain Imaging Data Structure}
}

\newglossaryentry{bedpostx}%
{
	name={BEDPOSTX},
	first={BEDPOSTX \cite[Bayesian Estimation of Diffusion Parameters Obtained using Sampling Techniques modelling crossing --X-- fibres,][]{jbabdi_modelbased_2012}},
	description={Bayesian Estimation of Diffusion Parameters Obtained using Sampling Techniques modelling crossing --X-- fibres}
}
\newacronym[first={FAST \cite[FMRIB's Automated Segmentation Tool,][]{zhang_segmentation_2001}}]%
{fast}{FAST}{FMRIB's Automated Segmentation Tool}
\newacronym[first={FIRST \cite[FMRIB's Integrated Registration and Segmentation Tool,][]{patenaude_bayesian_2011}}]%
{first}{FIRST}{FMRIB's Integrated Registration and Segmentation Tool}

### 2015-12-22 paper/main.tex
reply reviewers' comments

\documentclass[english]{frontiers/frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
\usepackage[mode=buildnew]{standalone}
\usepackage{tikz}
\usepackage[framemethod=TikZ]{mdframed}
%\setcitestyle{square}
\usepackage{url,lineno,microtype}
\usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
\usepackage[colorlinks=true,linkcolor=black, citecolor=black!80, urlcolor=black!80]{hyperref}
\usepackage{doi}
\usepackage[onehalfspacing]{setspace}

\def\keyFont{\fontsize{7}{9}\helveticabold }
\def\firstAuthorLast{Esteban {et~al.}} %use et al only if is more than 1 author
\def\Authors{Oscar Esteban\,$^{1,2,*}$, Emmanuel Caruyer\,$^{3}$, Alessandro Daducci\,$^{4}$, Meritxell Bach-Cuadra\,$^{4,5}$,%
Mar\'ia-J. Ledesma-Carbayo\,$^{1,2}$ and Andres Santos\,$^{1,2}$}

@@

\newcommand*\glslines[2][]{\glslink[#1]{#2}{\glsdisp{#2}{\acrlong{#2} --\glsentryshort{#2}--}}}

@@

\maketitle

\begin{mdframed}[hidealllines=true, backgroundcolor=black!60, roundcorner=10pt]
{\color{white} Data associated to this Report is available in the Dryad Digital Repository (\doi{10.5061/dryad.4p080}).}
\end{mdframed}

\linenumbers

@@

We recommend \diffantom{} as ground-truth in verification and validation frameworks
  (\autoref{fig:diffantom-02}, panel B) for testing pipelines.
\Diffantom{} is applicable in the unit testing of algorithms, the integration testing of
  modules in workflows, and the overall system testing.
Some potential applications follow:
\begin{itemize}
  \item Investigating the impact of different diffusion sampling schemes on the local microstructure
    model of choice and on the subsequent global tractography outcome.
  Since the gradient scheme can be set by the user, \diffantom{} can be seen as a mean to translate the so-called
  \emph{b-matrix} of the source dataset to any target scheme.
  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \gls*{csf} contamination,
    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and systems testing levels.
  \item Using \diffantom{} as in panel B of \autoref{fig:diffantom-02}, it is possible to apply binary classification measures
    to evaluate the resulting connectivity matrix.
  Considering the connectivity matrix of the \emph{reference Diffantom} and the resulting
    matrix of the \emph{test Diffantom}, the \acrfull*{roc} of the pipeline can be characterized.
  \item Simulation of pathological brains by altering the microstructural
    model accordingly (e.g. as tumors were simulated in \cite{kaus_simulation_2000}).
\end{itemize}
In order to exemplify one of these intended uses, we also release a \diffantom{} including the susceptibility-derived
  distortion in simulation.

>>>>

\documentclass[english]{frontiers/frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
\usepackage{xifthen}
\newboolean{review}
\setboolean{review}{true}

\usepackage[mode=buildnew]{standalone}
\usepackage{tikz}
\usepackage[framemethod=TikZ]{mdframed}
%\setcitestyle{square}
\usepackage{url,microtype}
\usepackage[right]{lineno}
\usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
\usepackage[colorlinks=true,linkcolor=black, citecolor=black!80, urlcolor=black!80]{hyperref}
\usepackage{doi}
\usepackage[onehalfspacing]{setspace}

\ifthenelse{\boolean{review}}
{
  \usepackage[colorinlistoftodos, textwidth=1.4cm]{todonotes}
  \setlength{\marginparwidth}{1.6cm}
  \reversemarginpar
  \newcommand{\revcomment}[2][]{\ifthenelse{\equal{#1}{}}{}{\todo[size=\tiny]{#1}}%
  \begingroup \color{blue!70} #2\endgroup}
}
{
  \newcommand{\todo}[2][]{}
  \newcommand{\revcomment}[2][]{#2}
}

% \usepackage{showframe}
\def\keyFont{\fontsize{7}{9}\helveticabold }
\def\firstAuthorLast{Esteban {et~al.}} %use et al only if is more than 1 author
\def\Authors{Oscar Esteban\,$^{1,2,*}$, Emmanuel Caruyer\,$^{3}$, Alessandro Daducci\,$^{4}$, Meritxell Bach-Cuadra\,$^{4,5}$,%
Mar\'ia-J. Ledesma-Carbayo\,$^{1,2}$ and Andres Santos\,$^{1,2}$}

@@

@@

\maketitle
\linenumbers

% \begin{mdframed}[hidealllines=true, backgroundcolor=black!5, roundcorner=5pt]
\section*{Diffantom in brief}
\revcomment[Rev.\#2\\Q.10]{%
\diffantom{} is a whole-brain \gls*{dmri} phantom publicly available
  through the Dryad Digital Repository (\doi{10.5061/dryad.4p080}).
The dataset contains two single-shell \gls*{dmri} images, along with the corresponding gradient
  information, packed following the \gls*{bids}.
The released dataset is designed for the evaluation of the impact of susceptibility distortions
  and benchmarking existing correction methods.
}%

\revcomment[Rev.\#1\\Q14.1]{%
In this Data Report we also release the software instruments involved in generating \emph{diffantoms},
  so that researchers are able to generate new phantoms derived from different subjects,
  and apply these data in other applications like investigating diffusion sampling
  schemes, the assessment of \gls*{dmri} processing methods, the simulation of pathologies
  and imaging artifacts, etc.
In summary, \diffantom{} is intended for unit testing of novel methods, cross-comparison of established methods,
  and integration testing of partial or complete processing flows to extract connectivity networks
  from \gls{dmri}.}

@@

We recommend \diffantom{} as ground-truth in verification and validation frameworks
  (\autoref{fig:diffantom-02}, panel B) for testing pipelines.
\Diffantom{} is applicable in the unit testing of algorithms, the integration testing of
  modules in workflows, and the overall system testing.
Some potential applications follow:
\begin{itemize}
  \item Investigating the impact of different diffusion sampling schemes on the local microstructure
    model of choice and on the subsequent global tractography outcome.
  Since the gradient scheme can be set by the user, \diffantom{} can be seen as a mean to translate the so-called
  \emph{b-matrix} of the source dataset to any target scheme.
  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \gls*{csf} contamination,
    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and systems testing levels.
  \item Using \diffantom{} as in panel B of \autoref{fig:diffantom-02}, it is possible to apply binary classification measures
    to evaluate the resulting connectivity matrix.
  Considering the connectivity matrix of the \emph{reference Diffantom} and the resulting
    matrix of the \emph{test Diffantom}, the \acrfull*{roc} of the pipeline can be characterized.
  \item Simulation of pathological brains by altering the microstructural
    model accordingly \cite[e.g. as tumors were simulated in][]{kaus_simulation_2000}.
\end{itemize}
In order to exemplify one of these intended uses, we also release a \diffantom{} including the susceptibility-derived
  distortion in simulation.
\revcomment[Rev.\#1\\Q14.2]{%
These two images belong to a broader dataset, automatically generated, used in a study to quantify
  the impact of susceptibility distortions and correction methods on the connectome extraction
  \cite[Chapter 5]{esteban_image_2015}.
In this study, three widely-used correction methods are compared in a reference framework of
  several \emph{Diffantoms} with realistic and controlled distortions.
This context provides a useful resource to characterize the impact of susceptibility distortion
  on the final connectivity network and allows the evaluation of the different correction
  methodologies available.%
}
